---
title: "Publication Data"
author: "Erica"
date: "2023-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression Analysis
```{r}
library(readxl)
library(dplyr)
library(devtools)
library(tidyverse)
library(ggbiplot)
library(ggplot2)
library(GGally)
library(mice) # for inputation
library(ggmice) # for inputation
library(scales)
#install.packages("corrplot")
library(corrplot)
library(writexl)

### Importing datasets
df <- read_excel("data_2000.xlsx")


```


### z-score function
```{r}

z_score <- function(x){
  zscore= ((x - mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE))
zscore
}

```


```{r}
df$employees_2001_norm <- z_score(df$employees_2001)
df$air_quality_norm <- z_score(df$air_quality)
df$water_quality_norm <- z_score(df$water_quality)
df$built_quality_norm <- z_score(df$built_quality)
df$land_quality_norm <- z_score(df$land_quality)
df$impervious_surface_norm <- z_score(df$impervious_surface)

```


##Plotting
## Causality_log and Property  Damage log
```{r}
################### causality  plot ################################
causality <- ggplot(df, aes(causality_log))
causality + geom_histogram(binwidth = 0.8)  ### left-skewed



################### Crop & Property Damage log ################################
crop_property_damage_log<- ggplot(df, aes(prop_dmg_log))
crop_property_damage_log + geom_histogram(binwidth = 1)

```

### CAUSALITY 
#### Social Dimension and Causality
```{r fig.height=10, fig.width=10}
### social dimension  VS causality log normalized plot
cor.test(df$causality_log, df$per_white_norm) #okay # remove due to highly correlation with black and per_female_hh_with_kids_under18_norm
cor.test(df$causality_log, df$per_black_norm)#okay
cor.test(df$causality_log, df$per_hispanic_norm)#okay
cor.test(df$causality_log, df$per_asian_norm)#okay
cor.test(df$causality_log, df$per_american_indian_norm)#okay 
cor.test(df$causality_log, df$per_other_races_norm)#okay# remove bc highly correlated with hispanci
cor.test(df$causality_log, df$per_POC_norm)#okay # remove bc highly correlated with hispanic

cor.test(df$causality_log, df$per_elderly_norm) #okay
cor.test(df$causality_log, df$per_young_dependent_norm)  #okay
cor.test(df$causality_log, df$per_noenglish_norm)# #okay highly correlated with hispanic
cor.test(df$causality_log, df$per_foreign_born_norm)##okay highly correlated with per_noenglish_norm
cor.test(df$causality_log, df$per_female_hh_with_kids_under6_norm)##okay
cor.test(df$causality_log, df$per_female_hh_with_kids_under18_norm)#okay # remove bc highly correlated with black, white, and per_female_hh_with_kids_under6_norm
cor.test(df$causality_log, df$per_rural_norm)#okay
cor.test(df$causality_log, df$per_no_schooling_completed_norm) #not statistically significant



X<-df%>%
  select(
         per_black_norm,
              per_hispanic_norm,
         per_asian_norm,
         per_american_indian_norm,
         per_elderly_norm,
         per_young_dependent_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm) 
        
ggpairs(X)


social_causality <- lm(causality_log~(per_black_norm+per_hispanic_norm + per_asian_norm + per_american_indian_norm+per_elderly_norm+per_young_dependent_norm+per_female_hh_with_kids_under6_norm+per_rural_norm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)
summary(social_causality)

```
#### Economic Dimension and Causality 
```{r fig.height=10, fig.width=10}
### economic dimension  VS causality log normalized plot

cor.test(df$causality_log, df$per_below_poverty_norm) #okay
cor.test(df$causality_log, df$median_hh_income_1999_norm)#okay # remove because highly correlated with per_below_poverty_norm and college 
cor.test(df$causality_log, df$per_rent_norm)# okay
cor.test(df$causality_log, df$per_no_carnorm)#okay

cor.test(df$causality_log, df$per_college_or_higher_norm)# okay
cor.test(df$causality_log, df$average_hh_norm)#okay

cor.test(df$causality_log, df$per_lack_plumbing_norm)#okay

cor.test(df$causality_log, df$per_lack_kitchen_norm) #okay # remove because highly correlated with lack of plumbing
cor.test(df$causality_log, df$per_mobile_home_norm)#okay
cor.test(df$causality_log, df$per_unemployed_norm)## not statistically significant


X<-df%>%
  select(
         per_below_poverty_norm,
         per_rent_norm,
         per_no_carnorm,
         per_college_or_higher_norm,
         per_lack_plumbing_norm,
         per_mobile_home_norm) 
        
ggpairs(X)


econ_causality <- lm(causality_log~(per_below_poverty_norm+
         per_rent_norm+
         per_no_carnorm+
         per_college_or_higher_norm+
         per_lack_plumbing_norm+
         per_mobile_home_norm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)
summary(econ_causality)

```

#### Health Dimension and Causality
```{r}
cor.test(df$causality_log, df$life_expectancy_2000_norm) # not okay
cor.test(df$causality_log, df$per_hypertension_2001_norm)# OKAY # remove due to high correlation
# cor.test(df$causality_log, df$per_heart_disease_35_65_norm)# not okay
# cor.test(df$causality_log, df$per_heart_disease_65_more_norm)# not okay
# cor.test(df$causality_log, df$per_stroke_35_65_norm) # OKAY
# cor.test(df$causality_log, df$per_stroke_65_more_norm) #not okay 
cor.test(df$causality_log, df$per_diabetes_2000_norm)# not okay
cor.test(df$causality_log, df$per_disability_norm)# OKAY
cor.test(df$causality_log, df$per_nursingnorm)# OKAY


X<-df%>%
  select(
         per_disability_norm,
         per_nursingnorm) 
        
ggpairs(X)

health_causality <- lm(causality_log~(
         per_disability_norm+
         per_nursingnorm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(health_causality)



```


#### Institutional Dimension and Causality
```{r}

cor.test(df$causality_log, df$FEMA_total_norm) # okay
cor.test(df$causality_log, df$number_research_institutions_norm) # okay
cor.test(df$causality_log, df$employees_2001_norm) # okay


X<-df%>%
  select(
         FEMA_total,
         number_research_institutions,
         employees_2001) 
        
ggpairs(X)

inst_causality <- lm(causality_log~(FEMA_total_norm+
         number_research_institutions_norm+
         employees_2001_norm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(inst_causality)




```


#### Environmental Dimension and Causality
```{r}

cor.test(df$causality_log, df$air_quality_norm)#okay
cor.test(df$causality_log, df$water_quality_norm)# not statistically significant
cor.test(df$causality_log, df$built_quality_norm)#okay
cor.test(df$causality_log, df$land_quality_norm)#not statistically significan
cor.test(df$causality_log, df$impervious_surface_norm)#okay

X<-df%>%
  select(
         air_quality_norm,
         built_quality_norm,
         impervious_surface_norm) 
        
ggpairs(X)

environ_causality <- lm(causality_log~(air_quality_norm+
         built_quality_norm+
         impervious_surface_norm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(environ_causality)

```

### Vulnerability using Causualties 
```{r model_all_causualties, fig.height=15, fig.width=15}


X<-df%>%
  select(per_black_norm,
              per_hispanic_norm,
         per_asian_norm,
         per_american_indian_norm,
         per_elderly_norm,
         per_young_dependent_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm,
         per_below_poverty_norm,
         per_rent_norm,
         per_no_carnorm,
         per_college_or_higher_norm,
         per_lack_plumbing_norm,
         per_mobile_home_norm,

         per_disability_norm,
         per_nursingnorm,
         FEMA_total_norm,
         number_research_institutions_norm,
         employees_2001_norm,
         air_quality_norm,
         built_quality_norm,
         impervious_surface_norm) 
        
ggpairs(X)


model_causality <- lm(causality_log~(per_black_norm+
              per_hispanic_norm+
         per_asian_norm+
         per_american_indian_norm+
         per_elderly_norm+
         per_young_dependent_norm+
         per_female_hh_with_kids_under6_norm+
         per_rural_norm+
         per_below_poverty_norm+
         per_rent_norm+
         per_no_carnorm+
         per_college_or_higher_norm+
         per_lack_plumbing_norm+
         per_mobile_home_norm+

         per_disability_norm+
         per_nursingnorm+
         FEMA_total_norm+
         number_research_institutions_norm+
         employees_2001_norm+
         air_quality_norm+
         built_quality_norm+
         impervious_surface_norm)+log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(model_causality)


```



### CROP and PROPERTY DAMAGE 
#### Social Dimension  and Prop Damage
```{r fig.height=10, fig.width=10}
### social dimension  VS crop & property dmg
cor.test(df$prop_dmg_log, df$per_white_norm) ## okay 
cor.test(df$prop_dmg_log, df$per_black_norm)# okay
cor.test(df$prop_dmg_log, df$per_hispanic_norm)##not statistically signif
cor.test(df$prop_dmg_log, df$per_asian_norm)#okay
cor.test(df$prop_dmg_log, df$per_american_indian_norm)# okay
cor.test(df$prop_dmg_log, df$per_other_races_norm)#not statistically signif
cor.test(df$prop_dmg_log, df$per_POC_norm)#okay


cor.test(df$prop_dmg_log, df$per_elderly_norm) ##okay
cor.test(df$prop_dmg_log, df$per_young_dependent_norm)  #okay
cor.test(df$prop_dmg_log, df$per_noenglish_norm)# ##not statistically signif
cor.test(df$prop_dmg_log, df$per_foreign_born_norm)##okay 
cor.test(df$prop_dmg_log, df$per_female_hh_with_kids_under6_norm)##okay
cor.test(df$prop_dmg_log, df$per_female_hh_with_kids_under18_norm)##okay - highly correlated
cor.test(df$prop_dmg_log, df$per_rural_norm)#okay
cor.test(df$prop_dmg_log, df$per_no_schooling_completed_norm) ##not statistically signif


X<-df%>%
  select(
         per_black_norm,
         per_asian_norm,
         per_american_indian_norm,
    per_elderly_norm,
         per_young_dependent_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm,
    per_foreign_born_norm) 
        
ggpairs(X)


social_damage <- lm(prop_dmg_log ~
                      per_black_norm+
         per_asian_norm+
           per_american_indian_norm+
    per_elderly_norm+
         per_young_dependent_norm+
    per_foreign_born_norm+
         per_female_hh_with_kids_under6_norm+
         per_rural_norm+
                        log_pop_2000+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(social_damage)



```
#### Economic Dimension  & prop damage 
```{r fig.height=10, fig.width=10}
### economic dimension  VS causality log normalized plot

cor.test(df$prop_dmg_log, df$per_below_poverty_norm) #okay
cor.test(df$prop_dmg_log, df$median_hh_income_1999_norm)##okay- highly correlated
cor.test(df$prop_dmg_log, df$per_rent_norm)# okay
cor.test(df$prop_dmg_log, df$per_no_carnorm)# okay

cor.test(df$prop_dmg_log, df$per_college_or_higher_norm)# okay
cor.test(df$prop_dmg_log, df$average_hh_norm)##okay

cor.test(df$prop_dmg_log, df$per_lack_plumbing_norm)##okay
cor.test(df$prop_dmg_log, df$per_lack_kitchen_norm) # okay - highly correlated
cor.test(df$prop_dmg_log, df$per_mobile_home_norm)# okay
cor.test(df$prop_dmg_log, df$per_unemployed_norm)## #not statistically significant


X<-df%>%
  select(
    per_below_poverty_norm, 
         per_rent_norm,
         per_no_carnorm,
         per_college_or_higher_norm,
    average_hh_norm,
    per_lack_plumbing_norm,
         per_mobile_home_norm) 
        
ggpairs(X)



econ_damage <- lm(prop_dmg_log ~per_below_poverty_norm+ 
         per_rent_norm+
         per_no_carnorm+
         per_college_or_higher_norm+
    average_hh_norm+
    per_lack_plumbing_norm+
         per_mobile_home_norm+
                        log_pop_2000+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(econ_damage)



```

#### Health Dimension  & prop damage
```{r}
cor.test(df$prop_dmg_log, df$life_expectancy_2000_norm) # okay- highly correlated
cor.test(df$prop_dmg_log, df$per_hypertension_2001_norm)# okay
cor.test(df$prop_dmg_log, df$per_heart_disease_35_65_norm)# okay- highly correlated
cor.test(df$prop_dmg_log, df$per_heart_disease_65_more_norm)# okay
cor.test(df$prop_dmg_log, df$per_stroke_35_65_norm) # okay- highly correlated
cor.test(df$prop_dmg_log, df$per_stroke_65_more_norm) #not statistically significant
cor.test(df$prop_dmg_log, df$per_diabetes_2000_norm)# okay - highly correlated
cor.test(df$prop_dmg_log, df$per_disability_norm)# not statistically significant
cor.test(df$prop_dmg_log, df$per_nursingnorm)# okay


X<-df%>%
  select(
    per_hypertension_2001_norm,
         per_nursingnorm) 
        
ggpairs(X)


health_damage <- lm(prop_dmg_log~(per_hypertension_2001_norm+
         per_nursingnorm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(health_damage)

df$predict_health_damage <- predict(health_damage, newdata = df)

```


#### Institutional Dimension  & prop damage
```{r}
cor.test(df$prop_dmg_log, df$FEMA_total_norm) # okay
cor.test(df$prop_dmg_log, df$number_research_institutions_norm) # okay
cor.test(df$prop_dmg_log, df$employees_2001_norm) # okay

X<-df%>%
  select(
         FEMA_total_norm,
         number_research_institutions_norm,
         employees_2001_norm) 
        
ggpairs(X)


inst_damage <- lm(prop_dmg_log~(FEMA_total_norm+
         number_research_institutions_norm+
         employees_2001_norm)+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(inst_damage)

```


#### Environmental Dimension  & prop damage
```{r}


cor.test(df$prop_dmg_log, df$air_quality_norm)#okay
cor.test(df$prop_dmg_log, df$water_quality_norm)# 
cor.test(df$prop_dmg_log, df$built_quality_norm)#okay
cor.test(df$prop_dmg_log, df$land_quality_norm)#not 
cor.test(df$prop_dmg_log, df$impervious_surface_norm)#okay

X<-df%>%
  select(
         air_quality_norm,
         water_quality_norm,
         built_quality_norm,
         land_quality_norm,
         impervious_surface_norm) 
        
ggpairs(X)

env_damage <- lm(prop_dmg_log~(air_quality_norm+
         water_quality_norm+
         built_quality_norm+
           land_quality_norm+
           impervious_surface_norm
           )+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(env_damage)



```

### Vulnerability using crop & prop damage 
```{r , fig.height=17, fig.width=17}

X<-df%>%
  select(
         per_black_norm,
         per_asian_norm,
    per_elderly_norm,
    per_noenglish_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm,
    per_below_poverty_norm, 
         per_rent_norm,
         per_no_carnorm,
         per_college_or_higher_norm,
    average_hh_norm,
    per_lack_plumbing_norm,
         per_mobile_home_norm,
    air_quality_norm,
         water_quality_norm,
         built_quality_norm,
         land_quality_norm,
         impervious_surface_norm,
         per_nursingnorm,
     FEMA_total_norm,
         number_research_institutions_norm,
         employees_2001_norm) 
        
ggpairs(X)







model_dmg <- lm(prop_dmg_log~ (per_black_norm+
         per_asian_norm+
    per_elderly_norm+
    per_noenglish_norm+
         per_female_hh_with_kids_under6_norm+
         per_rural_norm)+
                        
                        (per_nursingnorm)+
                        
                        (per_below_poverty_norm+ 
         per_rent_norm+
         per_no_carnorm+
         per_college_or_higher_norm+
    average_hh_norm+
    per_lack_plumbing_norm+
         per_mobile_home_norm)
                      
                        +  
                (FEMA_total_norm+
         number_research_institutions_norm+
         employees_2001_norm)        
                        +
          (air_quality_norm+
         water_quality_norm+
         built_quality_norm+
           land_quality_norm+
           impervious_surface_norm )+
                        log_pop_2000+numb_haz_log+state,data=df, na.rm=TRUE)

summary(model_dmg)

```


