---
title: "NHV 2010"
author: "Erica"
date: "2023-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression analysis
```{r}

library(readxl)
library(dplyr)
library(devtools)
library(tidyverse)
library(ggplot2)
library(GGally)
library(writexl)

df <- read_excel("data_2010.xlsx")

```


### z-score function
```{r}

z_score <- function(x){
  zscore=(x - mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE)
zscore
}

```


##Plotting
## Causality_log and Property  Damage log
```{r}
################### causality  plot ################################
causality <- ggplot(df, aes(causality_log))
causality + geom_histogram(binwidth = 0.8)  ### left-skewed  



################### Property Damage log ################################
property_damage_log<- ggplot(df, aes(prop_dmg_log))
property_damage_log + geom_histogram(binwidth = 1)  

```

### CAUSALITY 
#### Social Dimension and Causality
```{r fig.height=10, fig.width=10}
### social dimension  VS causality log normalized plot
cor.test(df$causality_log, df$per_white_norm) #okay
cor.test(df$causality_log, df$per_black_norm)#okay
cor.test(df$causality_log, df$per_hispanic_alone_norm)#okay
cor.test(df$causality_log, df$per_asian_norm)# okay
cor.test(df$causality_log, df$per_american_indian_norm)# no
cor.test(df$causality_log, df$per_other_races_norm)#okay
cor.test(df$causality_log, df$per_POC_norm)#okay


df$per_hispanic_norm <- df$per_hispanic_alone_norm # since i had it with the wrong name

cor.test(df$causality_log, df$per_elderly_norm) # okay
cor.test(df$causality_log, df$per_young_dependent_norm)  #okay
cor.test(df$causality_log, df$per_noenglish_norm)# #okay but removed due to correlation with hisp and foreign born
cor.test(df$causality_log, df$per_foreign_born_norm)##okay
cor.test(df$causality_log, df$per_female_hh_with_kids_under6_norm)#okay
cor.test(df$causality_log, df$per_female_hh_with_kids_under18_norm)#okay
cor.test(df$causality_log, df$per_rural_norm)#okay
cor.test(df$causality_log, df$per_per_no_school_completed_norm) #not statistically significant


X<-df%>%
  select(
         per_black_norm,
              per_hispanic_norm,
                       per_asian_norm,
         per_elderly_norm,
         per_young_dependent_norm,
         per_foreign_born_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm) 
        
ggpairs(X)


social_causality <- lm(causality_log~(per_black_norm+per_hispanic_norm + per_asian_norm + per_elderly_norm+per_young_dependent_norm+per_foreign_born_norm+per_female_hh_with_kids_under6_norm+per_rural_norm)+
                       log_pop_2010+numb_haz_log+state,data=df, na.rm=TRUE)
summary(social_causality)


```
#### Economic Dimension and Causality 
```{r fig.height=10, fig.width=10}
### economic dimension  VS causality log normalized plot

cor.test(df$causality_log, df$per_below_poverty_norm) # not statistically significant
cor.test(df$causality_log, df$median_hh_income_2010_norm)# okay
cor.test(df$causality_log, df$per_rent_norm)# okay but removed do to correlation with no car
cor.test(df$causality_log, df$per_no_carnorm)#okay

cor.test(df$causality_log, df$per_college_or_higher_norm)# okay but removed due to correlation with hh income
cor.test(df$causality_log, df$average_hh_norm)#okay

cor.test(df$causality_log, df$per_lack_plumbing_norm)#okay

cor.test(df$causality_log, df$per_lack_kitchen_norm) #okay
cor.test(df$causality_log, df$per_mobile_home_norm)#okay
cor.test(df$causality_log, df$per_unemployed_norm)## okay

X<-df%>%
  select(
         per_no_carnorm,
         median_hh_income_2010_norm,
         average_hh_norm,
         per_lack_plumbing_norm,
         per_lack_kitchen_norm,
         per_mobile_home_norm,
         per_unemployed_norm) 
        
ggpairs(X)


econ_causality <- lm(causality_log~(
         per_no_carnorm+median_hh_income_2010_norm+average_hh_norm+
         per_lack_plumbing_norm+per_lack_kitchen_norm+
         per_mobile_home_norm+per_unemployed_norm)+
                        log_pop_2010+numb_haz_log+state,data=df, na.rm=TRUE)
summary(econ_causality)

```

#### Health Dimension and Causality
```{r}
cor.test(df$causality_log, df$life_expectancy_2010_norm) # okay
cor.test(df$causality_log, df$per_hypertension_2009_norm)# no
cor.test(df$causality_log, df$per_diabetes_2010_norm)#okay removed due to high correlation
cor.test(df$causality_log, df$per_disability_norm)# okay
cor.test(df$causality_log, df$per_nursingnorm)# okay
#cor.test(df$causality_norm_log, df$per_institutionalized_norm)## OKAY



X<-df%>%
  select(life_expectancy_2010_norm,
            per_disability_norm,
         per_nursingnorm) 
        
ggpairs(X)

health_causality <- lm(causality_log~(life_expectancy_2010_norm+
         per_disability_norm+
         per_nursingnorm)+
                        log_pop_2010+numb_haz_norm+state,data=df, na.rm=TRUE)

summary(health_causality)



```


#### Institutional Dimension and Causality
```{r}
cor.test(df$causality_log, df$FEMA_total_norm) # okay
cor.test(df$causality_log, df$number_research_institutions_norm) # okay
cor.test(df$causality_log, df$employees_2010_norm) # okay


X<-df%>%
  select(
         FEMA_total_norm,
         number_research_institutions_norm,
         employees_2010_norm) 
        
ggpairs(X)

inst_causality <- lm(causality_log~(FEMA_total_norm+
         number_research_institutions_norm+
         employees_2010_norm)+
                        log_pop_2010+numb_haz_log+state,data=df, na.rm=TRUE)

summary(inst_causality)

```


#### Environmental Dimension and Causality
```{r}
df$air_quality_norm <- z_score(df$air_quality)
df$water_quality_norm <- z_score(df$water_quality)
df$built_quality_norm <- z_score(df$built_quality)
df$land_quality_norm <- z_score(df$land_quality)



cor.test(df$causality_log, df$air_quality_norm)# okay
cor.test(df$causality_log, df$water_quality_norm)# no
cor.test(df$causality_log, df$built_quality_norm)#okay
cor.test(df$causality_log, df$land_quality_norm)#okay
cor.test(df$causality_log, df$impervious_surface_norm)#okay


X<-df%>%
  select(
         air_quality_norm,
         built_quality_norm,
         land_quality_norm,
         impervious_surface_norm) 
        
ggpairs(X)

environ_causality <- lm(causality_log~(air_quality_norm+
         built_quality_norm+
           land_quality_norm+
         impervious_surface_norm)+
                        log_pop_2010+numb_haz_log+state,data=df, na.rm=TRUE)

summary(environ_causality)

```

### Aggregated Casualties
```{r fig.height=15, fig.width=15}
X<-df%>%
  select(
         per_black_norm,
              per_hispanic_norm,
                       per_asian_norm,
         per_elderly_norm,
         per_young_dependent_norm,
         per_foreign_born_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm,
         
         per_no_carnorm,
         median_hh_income_2010_norm,
         average_hh_norm,
         per_lack_plumbing_norm,
         per_lack_kitchen_norm,
         per_mobile_home_norm,
         per_unemployed_norm,
         
         life_expectancy_2010_norm,
         per_disability_norm,
         per_nursingnorm,
         
         FEMA_total_norm,
         number_research_institutions_norm,
         employees_2010_norm,
         
         air_quality_norm,
         built_quality_norm,
         land_quality_norm,
         impervious_surface_norm
         
         ) 
        
ggpairs(X)




aggregated_causality <- lm(causality_log~(per_black_norm+
              per_hispanic_norm+
                       per_asian_norm+
         per_elderly_norm+
         per_young_dependent_norm+
         per_foreign_born_norm+
         per_female_hh_with_kids_under6_norm+
         per_rural_norm+
         
         per_no_carnorm+
         median_hh_income_2010_norm+
         per_lack_plumbing_norm+
         per_lack_kitchen_norm+
         per_mobile_home_norm+
         per_unemployed_norm+
         
         life_expectancy_2010_norm+
         per_disability_norm+
         per_nursingnorm+
         
         FEMA_total_norm+
         number_research_institutions_norm+
         employees_2010_norm+
         
         air_quality_norm+
         built_quality_norm+
         land_quality_norm+
         impervious_surface_norm)+
                        log_pop_2010+numb_haz_log+state,data=df, na.rm=TRUE)

summary(aggregated_causality)



```



### CROP and PROPERTY DAMAGE 
#### Social Dimension  and Prop Damage
```{r fig.height=10, fig.width=10}
### social dimension  VS property dmg
cor.test(df$prop_dmg_log, df$per_white_norm) ## okay
cor.test(df$prop_dmg_log, df$per_black_norm)# okay
cor.test(df$prop_dmg_log, df$per_hispanic_norm)##not statistically signif
cor.test(df$prop_dmg_log, df$per_asian_norm)#okay
cor.test(df$prop_dmg_log, df$per_american_indian_norm)# okay
cor.test(df$prop_dmg_log, df$per_other_races_norm)#okay remove due to correlation
cor.test(df$prop_dmg_log, df$per_POC_norm)#okay

cor.test(df$prop_dmg_log, df$per_elderly_norm) ##okay
cor.test(df$prop_dmg_log, df$per_young_dependent_norm)  #okay
cor.test(df$prop_dmg_log, df$per_noenglish_norm)# ##okay remove due to correlation
cor.test(df$prop_dmg_log, df$per_foreign_born_norm)##okay
cor.test(df$prop_dmg_log, df$per_female_hh_with_kids_under6_norm)## okay
cor.test(df$prop_dmg_log, df$per_female_hh_with_kids_under18_norm)##okay remove due to correlation
cor.test(df$prop_dmg_log, df$per_rural_norm)#okay
cor.test(df$prop_dmg_log, df$per_per_no_school_completed_norm) ## okay


X<-df%>%
  select(              
         per_asian_norm,
         per_black_norm,
         per_elderly_norm,
         per_american_indian_norm,
         per_young_dependent_norm,
         per_foreign_born_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm,
         per_per_no_school_completed_norm) 
        

ggpairs(X)


social_dmg <- lm(prop_dmg_log~(
         per_asian_norm+
         per_black_norm+
         per_elderly_norm+
         per_american_indian_norm+
         per_young_dependent_norm+
         per_foreign_born_norm+
         per_female_hh_with_kids_under6_norm+
         per_rural_norm+
         per_per_no_school_completed_norm)+
                        log_pop_2010+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(social_dmg)



```

#### Economic Dimension and damage 
```{r fig.height=10, fig.width=10}

cor.test(df$prop_dmg_log, df$per_below_poverty_norm) # no
cor.test(df$prop_dmg_log, df$median_hh_income_2010_norm)##
cor.test(df$prop_dmg_log, df$per_rent_norm)# 
cor.test(df$prop_dmg_log, df$per_no_carnorm)# removed bc highly correlated with rent

cor.test(df$prop_dmg_log, df$per_college_or_higher_norm)#
cor.test(df$prop_dmg_log, df$average_hh_norm)##

cor.test(df$prop_dmg_log, df$per_lack_plumbing_norm)## no
cor.test(df$prop_dmg_log, df$per_lack_kitchen_norm) #
cor.test(df$prop_dmg_log, df$per_mobile_home_norm)#no
cor.test(df$prop_dmg_log, df$per_unemployed_norm)## 


X<-df%>%
  select(median_hh_income_2010_norm,
         per_rent_norm,
         average_hh_norm,
         per_college_or_higher_norm,
         per_lack_kitchen_norm,
         per_unemployed_norm) 
        
ggpairs(X)


econ_dmg <- lm(prop_dmg_log~(median_hh_income_2010_norm+
         per_rent_norm+
         average_hh_norm+
         per_college_or_higher_norm+
         per_lack_kitchen_norm+
         per_unemployed_norm)+
                        log_pop_2010+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(econ_dmg)

```

#### Health Dimension and damage
```{r}
cor.test(df$prop_dmg_log, df$life_expectancy_2010_norm) # highly correlated
cor.test(df$prop_dmg_log, df$per_hypertension_2009_norm)# highly correlated
cor.test(df$prop_dmg_log, df$per_heart_disease_35_65_norm)# highly correlated
cor.test(df$prop_dmg_log, df$per_heart_disease_65_more_norm)# 
cor.test(df$prop_dmg_log, df$per_stroke_35_65_norm) # 
cor.test(df$prop_dmg_log, df$per_stroke_65_more_norm) #no
cor.test(df$prop_dmg_log, df$per_diabetes_2010_norm)# 
cor.test(df$prop_dmg_log, df$per_disability_norm)# no
cor.test(df$prop_dmg_log, df$per_nursingnorm)# okay


X<-df%>%
  select(
         per_diabetes_2010_norm,
         per_nursingnorm) 
        
ggpairs(X)


health_dmg <- lm(prop_dmg_log~(per_diabetes_2010_norm+
         per_nursingnorm)+
                        log_pop_2010+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(health_dmg)

```


#### Institutional Dimension and damage
```{r}
cor.test(df$prop_dmg_log, df$FEMA_total_norm) # okay
cor.test(df$prop_dmg_log, df$number_research_institutions_norm) # okay
cor.test(df$prop_dmg_log, df$employees_2010_norm) # okay

X<-df%>%
  select(
         FEMA_total_norm,
         number_research_institutions_norm,
         employees_2010_norm) 
        
ggpairs(X)


inst_dmg <- lm(prop_dmg_log~(FEMA_total_norm+
         number_research_institutions_norm+
         employees_2010_norm)+
                        log_pop_2010+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(inst_dmg)


```


#### Environmental Dimension and damage
```{r}
df$air_quality_norm <- z_score(df$air_quality)
df$water_quality_norm <- z_score(df$water_quality)
df$built_quality_norm <- z_score(df$built_quality)
df$land_quality_norm <- z_score(df$land_quality)



cor.test(df$prop_dmg_log, df$air_quality_norm)#okay 
cor.test(df$prop_dmg_log, df$water_quality_norm)# no
cor.test(df$prop_dmg_log, df$built_quality_norm)#okay
cor.test(df$prop_dmg_log, df$land_quality_norm)#
cor.test(df$prop_dmg_log, df$impervious_surface_norm)#okay

X<-df%>%
  select(air_quality_norm,
         built_quality_norm,
         land_quality_norm,
         impervious_surface_norm) 
        
ggpairs(X)


env_dmg <- lm(prop_dmg_log~(air_quality_norm+
         built_quality_norm+
         land_quality_norm+
         impervious_surface_norm)+
                        log_pop_2010+numb_haz_log+state+log_median_house_value,data=df, na.rm=TRUE)

summary(env_dmg)


```

### Vulnerability using Dmg 
```{r fig.height=17, fig.width=17}

X<-df%>%
  select(
        per_asian_norm,
         per_black_norm,
         per_elderly_norm,
         per_american_indian_norm,
         per_young_dependent_norm,
         per_foreign_born_norm,
         per_female_hh_with_kids_under6_norm,
         per_rural_norm,
         per_per_no_school_completed_norm,
         median_hh_income_2010_norm,
         per_rent_norm,
         average_hh_norm,
         per_college_or_higher_norm,
         per_lack_kitchen_norm,
         per_unemployed_norm,
         per_diabetes_2010_norm,
         per_nursingnorm,
                        FEMA_total_norm,
         number_research_institutions_norm,
         employees_2010_norm,
                        
                    air_quality_norm,
         built_quality_norm,
         land_quality_norm,
         impervious_surface_norm
         
         
         ) 
        
ggpairs(X)






model_dmg <- lm(prop_dmg_log~
                        (per_asian_norm + 
    per_black_norm + per_elderly_norm + per_american_indian_norm + 
    per_young_dependent_norm + per_foreign_born_norm + per_female_hh_with_kids_under6_norm + 
    per_rural_norm + per_per_no_school_completed_norm)+
                        
                        (median_hh_income_2010_norm+
         per_rent_norm+
           average_hh_norm+
         per_college_or_higher_norm+
         per_lack_kitchen_norm+
         per_unemployed_norm)+
                        (per_diabetes_2010_norm+
         per_nursingnorm)+
                        
                        (FEMA_total_norm+
         number_research_institutions_norm+
         employees_2010_norm)+
                    (air_quality_norm+
         built_quality_norm+
         land_quality_norm+
         impervious_surface_norm)+
                        log_pop_2010+numb_haz_log+state+log_median_house_value    
                        
                        
                        
                        ,data=df, na.rm=TRUE)


summary(model_dmg)

```


